Applications of limited information strategies in Menger's game

### 2015-09-23 articles/applications-of-limited-information-strategies-in-mengers-game/article.tex
fix section title
polish paper, missing citation for 4.3 and intro/abstr

Note that many authors refer to this property as \(S_{fin}(\mc O,\mc O)\),
where \(\mc O\) is the collection of open covers of \(X\), and
\(S_{fin}(A,B)\) denotes the selection property such that for each
sequence in \(A^\omega\), there are finite subsets of each entry for which
the union of these subsets belongs in \(B\).

@@

As with the Menger property, authors usually characterize this game using
finite subsets \(\mc F_n\)
of \(\mc U_n\) instead (and often refer to it as \(G(\mc O,\mc O)\) as
with the selection property above).
This is obviously equivalent in the case of perfect
information, and is also equivalent
in the case of limited information, provided \(\pl F\)
knows \(\mc U_n\) during round \(n\). However, we make this change as we
will investigate \(0\)-Markov strategies which consider no moves of
the opponent, and instead rely only on the current round number.

@@

@@

\begin{proposition}
  \(\pl F \notmarkwin \menGame{\oneptlind{\omega_1}}\); in fact,
  \(\oneptlind{\omega_1}\) is not \(\sigma\)-compact.
\end{proposition}

The greatest advantage of a strategy which has knowledge of two or more previous
moves of the opponent, versus only knowledge of the most recent move, is the
ability to react to changes from one round to the next. It's this ability to
react that will give \(\pl F\) her winning \(2\)-Markov strategy in the Menger
game on \(\oneptlind\omega_1\). This may be proven directly; however, we will
postpone giving a proof so that we may take advantage of an equivalent
game characterization of this property.

@@

While we have not shown a direct implication between the Menger game and
Scheeper's original filling game, Scheepers introduced the statement
\(\alcompS\kappa\) relating to the almost-compatability of functions
from countable subsets of \(\kappa\) into \(\omega\) which may be applied to
both.

@@

\begin{definition}
  \(\alcompS\kappa\) states that there exist functions
  \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\omega}\) such that
  \(|\{\alpha\in A:f_A(\alpha)\leq n\}|<\omega\) for all \(n<\omega\) and
  \(f_A\alcomp f_B\) for all \(A,B\in[\kappa]^\omega\).
  \footnote{
  This is equivalent to the original characterization given in
  \cite{MR1129143}: there exist injections \(g_A:A\to\omega\)
  such that \(g_A\alcomp g_B\) for all \(A,B\in[\kappa]^\omega\) and \(A\subset B\).
  }
\end{definition}

@@

     \text{Markov Menger} &
     2\text{-Markov Menger} &
     \text{Perfect Menger} &
     \text{Menger} \\
  };
  \path[>=latex,<->]
    (m-1-1) edge (m-2-1)
    (m-1-2) edge (m-2-2)
    (m-1-3) edge (m-2-3)
    (m-1-4) edge (m-2-4);
  \path[>=latex,->]
    (m-1-1) edge (m-1-2)
    (m-2-1) edge (m-2-2)
    (m-1-2) edge (m-1-3)
    (m-2-2) edge (m-2-3)
    (m-1-3) edge (m-1-4)
    (m-2-3) edge (m-2-4);
  \path[>=latex,->,bend right=30, dotted]
    (m-1-2) edge node {\( \not \)} (m-1-1)
    (m-2-2) edge node {\( \not \)} (m-2-1)
    (m-1-4) edge node {\( \not \)} (m-1-3)
    (m-2-4) edge node {\( \not \)} (m-2-3);
  \path[>=latex,->,bend right=30, dotted]
    (m-1-3) edge node {\( ? \)} (m-1-2)
    (m-2-3) edge node {\( ? \)} (m-2-2);
\end{tikzpicture}
\end{center}
\caption{Diagram of covering properties related to the Menger game}
\label{menSpec}
\end{figure}

@@

We are also interested in non-game-theoretic characterizations of these
covering properties. It has been known for some time that metrizable perfect
Menger spaces are exactly the metrizable \(\sigma\)-compact spaces, shown first
by Telgarksy in \cite{MR753073} and later directly by Scheepers in
\cite{MR1273523}.

@@

\begin{lemma}
  Let \(\sigma(\mc{U}, n)\) be a Markov strategy for \(F\) in
  \(\menGame{X}\), and \(\mathfrak{C}\) collect all open covers of \(X\). Then the
  set
    \[
      R_n = \bigcap_{\mc{U}\in\mathfrak{C}} \sigma(\mc{U},n)
    \]
  is relatively compact to \(X\). If \(\sigma\) is a winning
  Markov startegy, then \(\bigcup_{n<\omega} R_n = X\).
\end{lemma}

\begin{proof}
  First, for every open cover \(\mc U\in\mathfrak C\),
  \(R_n\subseteq\sigma(\mc U,n)\) is covered by a finite subcollection of \(\mc U\).

  Suppose that \(x \not\in R_n\) for any \(n<\omega\). Then for each \(n\),
  pick \(\mc{U}_n\in\mathfrak{C}\) such that \(x\not\in \sigma(\mc{U}_n,n)\). Then
  \(\pl C\) may counter \(\sigma\) with the attack \(\<\mc U_0,\mc U_1,\dots\>\).
\end{proof}

@@

For Lindel\"of spaces, metrizability is characterized by regularity and
second-countability, the latter of which was essentially used by Scheepers in
this way:

@@

  Let \(t:\omega\to\omega^{<\omega}\) be a bijection. We define the Mark\"ov
  strategy \(\tau\) as follows:
  \[
    \tau(\<\mc V\>, n)
      =
    \bigcup \mc F(t(n),\mc V)
  \]

@@

\section{Robustly Lindel\"of}

@@

We will soon see that, assuming \(\alcompS{\mf c}\), \(\pl F\) has a winning
\(2\)-Mark\"ov strategy for \(\menGame{R_\omega}\) as well.

\begin{proposition}
  If \(X=\bigcup_{i<\omega} X_i\) and \(\pl F \kmarkwin{k} \menGame{X,X_i}\)
  for \(i<\omega\), then \(\pl F \kmarkwin{k} \menGame{X}\)
\end{proposition}

\begin{proof}
  Let \(\sigma_i\) be a \(k\)-Markov strategy
  for \(\pl F\) in \(\menGame{X,X_i}\).

  We define the \(k\)-Markov strategy
  \(\sigma\) for \(\menGame{X}\) as follows:
    \[
      \sigma(\<\mc U_0,\dots,\mc U_n\>)
        =
      \bigcup_{i\leq n}
      \sigma_i(\<\mc U_i,\dots,\mc U_n\>)
    \]

  Let \(\<\mc U_0,\mc U_1,\dots\>\) be a successful counter-attack by \(\pl C\)
  against \(\sigma\). Then there exists \(x\in X_i\) for
  some \(i<\omega\) such that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma(\<\mc U_0,\dots,\mc U_n\>)\).
  It follows that \(x\) is not covered by
  \(\bigcup_{n<\omega}\sigma_i(\<\mc U_i,\dots,\mc U_{i+n}\>)\),
  and \(\<\mc U_i,\mc U_{i+1},\dots\>\) is a successful counter-attack
  by \(\pl C\) against \(\sigma_i\).
\end{proof}

@@

>>>>

Note that many authors refer to this property as \(S_{fin}(\mc O,\mc O)\)
\cite{MR1378387},
where \(\mc O\) is the collection of open covers of \(X\), and
\(S_{fin}(A,B)\) denotes the selection property such that for each
sequence in \(A^\omega\), there are finite subsets of each entry for which
the union of these subsets belongs in \(B\).

@@

As with the Menger property, authors usually characterize this game using
finite subsets \(\mc F_n\)
of \(\mc U_n\) instead (and often refer to it as the selection game
\(G_{fin}(\mc O,\mc O)\)).
This is obviously equivalent in the case of perfect
information, and is also equivalent
in the case of limited information, provided \(\pl F\)
knows \(\mc U_n\) during round \(n\). However, we make this change as we
will investigate \(0\)-Markov strategies which consider no moves of
the opponent, and instead rely only on the current round number.

@@

It's worth noting that the above proof holds for any selection game of
the form \(G_{fin}(A,B)\).

@@

One may prove directly that
  \(\pl F \notmarkwin \menGame{\oneptlind{\omega_1}}\)
without much effort; however, we postpone stating this until we have
shown that being Markov Menger is equivalent to being \(\sigma\)-compact
in regular spaces.

The greatest advantage of a strategy which has knowledge of two or more previous
moves of the opponent, versus only knowledge of the most recent move, is the
ability to react to changes from one round to the next. It's this ability to
react that often allows a player to wield a winning \(2\)-Markov strategy when
a winning \(1\)-Markov strategy does not exist.

@@

While we have not shown a direct implication between the Menger game and
Scheeper's original filling game, Scheepers introduced the statement
\(\alcompS\kappa\) relating to the almost-compatability of injective functions
from countable subsets of \(\kappa\) into \(\omega\) which may be applied to
both.

@@

\begin{definition}
  \(\alcompS\kappa\) states that there exist injective functions
  \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\omega}\) such that
  \(f_A\alcomp f_B\) for all \(A,B\in[\kappa]^\omega\).
\end{definition}

@@

     \text{Markov Menger} &
     2\text{-Markov Menger} &
     \text{Strategic Menger} &
     \text{Menger} \\
  };
  \path[>=latex,<->]
    (m-1-1) edge (m-2-1)
    (m-1-2) edge (m-2-2)
    (m-1-3) edge (m-2-3)
    (m-1-4) edge (m-2-4);
  \path[>=latex,->]
    (m-1-1) edge (m-1-2)
    (m-2-1) edge (m-2-2)
    (m-1-2) edge (m-1-3)
    (m-2-2) edge (m-2-3)
    (m-1-3) edge (m-1-4)
    (m-2-3) edge (m-2-4);
  \path[>=latex,->,bend right=30, dotted]
    (m-1-2) edge node {\( \not \)} (m-1-1)
    (m-2-2) edge node {\( \not \)} (m-2-1)
    (m-1-4) edge node {\( \not \)} (m-1-3)
    (m-2-4) edge node {\( \not \)} (m-2-3);
  \path[>=latex,->,bend right=30, dotted]
    (m-1-3) edge node {\( ? \)} (m-1-2)
    (m-2-3) edge node {\( ? \)} (m-2-2);
\end{tikzpicture}
\end{center}
\caption{Diagram of covering properties related to the Menger game}
\label{menSpec}
\end{figure}

@@

We are also interested in non-game-theoretic characterizations of these
covering properties. It has been known for some time that metrizable perfect
Menger spaces are exactly the metrizable \(\sigma\)-compact spaces, shown first
by Telgarksy in \cite{MR753073} and later directly by Scheepers in
\cite{MR1273523}. It's Scheepers' technique which gives inspiration to
the following.

@@

%TODO cite source?

\begin{lemma}
  Let \(\sigma(\mc{U}, n)\) be a Markov strategy for \(F\) in
  \(\menGame{X}\), and \(\mc{O}\) collect all open covers of \(X\). Then the
  set
    \[
      R_n = \bigcap_{\mc{U}\in\mc{O}} \sigma(\mc{U},n)
    \]
  is relatively compact to \(X\). If \(\sigma\) is a winning
  Markov startegy, then \(\bigcup_{n<\omega} R_n = X\).
\end{lemma}

\begin{proof}
  First, for every open cover \(\mc U\in\mc{O}\),
  \(R_n\subseteq\sigma(\mc U,n)\) is covered by a finite subcollection of \(\mc U\).

  Suppose that \(x \not\in R_n\) for any \(n<\omega\). Then for each \(n\),
  pick \(\mc{U}_n\in\mc{O}\) such that \(x\not\in \sigma(\mc{U}_n,n)\). Then
  \(\pl C\) may counter \(\sigma\) with the attack \(\<\mc U_0,\mc U_1,\dots\>\).
\end{proof}

@@

\begin{corollary}
  \(\pl F\notmarkwin\menGame{\oneptlind{\omega_1}}\).
\end{corollary}

Note that for Lindel\"of spaces, metrizability is characterized by regularity
and second-countability.

@@

  Let \(t:\omega\to\omega^{<\omega}\) be a bijection. We define the Markov
  strategy \(\tau\) as follows:
  \[
    \tau(\<\mc V\>, n)
      =
    \bigcup \mc F(t(n),\mc V)
  \]

@@

\section{Robustly Menger}

@@

We will soon see that, assuming \(\alcompS{\mf c}\), \(\pl F\) has a winning
\(2\)-Markov strategy for \(\menGame{R_\omega}\) as well.

\begin{proposition}
  If \(X=\bigcup_{i<\omega} X_i\) and \(\pl F \kmarkwin{2} \menGame{X,X_i}\)
  for \(i<\omega\), then \(\pl F \kmarkwin{2} \menGame{X}\)
\end{proposition}

\begin{proof}
  Let \(\sigma_i\) be a \(2\)-Markov strategy
  for \(\pl F\) in \(\menGame{X,X_i}\).

  We define the \(2\)-Markov strategy
  \(\sigma\) for \(\menGame{X}\) as follows:
    \[
      \sigma(\<\mc U\>,0) = \sigma_0(\<\mc U\>,0)
    \]
    \[
      \sigma(\<\mc U,\mc V\>,n+1)
        =
      \bigcup_{i,j\leq n}
      \sigma_i(\<\mc V\>,0)
        \cup
      \sigma_i(\<\mc U,\mc V\>,j+1)
    \]

  Let \(\<\mc U_0,\mc U_1,\dots\>\) be a successful counter-attack by \(\pl C\)
  against \(\sigma\). Then there exists \(x\in X_i\) for
  some \(i<\omega\) such that \(x\) is not contained in
  \[
    \sigma(\<\mc U_0\>,0)
      \cup
    \bigcup_{n<\omega}
    \sigma(\<\mc U_n,\mc U_{n+1}\>,n+1)
  \]
  It then follows that \(x\) is not contained in
  \[
    \sigma_i(\<\mc U_i\>,0)
      \cup
    \bigcup_{n<\omega}
    \sigma_i(\<\mc U_{i+n},\mc U_{i+n+1}\>,n+1)
  \]
  and \(\<\mc U_i,\mc U_{i+1},\dots\>\) is a successful counter-attack
  by \(\pl C\) against \(\sigma_i\).
\end{proof}

@@

The following is left open:

\begin{question}
  Does there exist a space which is \(2\)-Markov Menger, but not
  robustly Menger?
\end{question}

\section*{Acknowlegements}

The author wishes to thank Gary Gruenhage and Alan Dow for their
very useful comments during the writing of this paper.
